Browsernative Microservices

### 2017-02-21 build/paper.tex
pdf/tex build

``A~\textbf{component}~is a unit of software that is independently
replaceable and upgradeable.'' {[}4{]} Components are the building
blocks of microservices. And microservices are the building blocks of
applications. Essentially the difference between microservices and
components is just the level of abstraction. Whether a concrete
microservice or a much more generic component, both following pretty
much the same principles. Therefore this paper referring to both parts
when talking about \textbf{services}.

The first principle of services is the \textbf{loose coupling
principle}: changing and deploying one service shouldn't result in
changing other parts of the system.{[}3, p. 30{]}. Mutations or shadowed
variables, which is happening a lot in CSS, making it hard to keep
changes ought to only affect one place in the application. A
\emph{browsernative microservice} therefore pushing encapsulation and
avoiding variable mutations outside its scope as much as possible.
Practically, CSS will be scoped and JS fosters immutable JS entities and
avoids variables leaking into the global namespace preferring the
\texttt{let} or \texttt{const} declaration in favor of the old
\texttt{var} declaration.

The second principle of services is the \textbf{high cohesion
principle}: Whether designing a microservice or it's components we want
related behavior sit together, and unrelated behavior to sit
elsewhere.{[}3, p. 30{]} High cohesion can be expressed in a dynamic way
as the \emph{Single Responsibility Principle}: ``Gather together those
things that change for the same reason and separate those things that
change for different reasons.''{[}5{]} In a very quick and dirty code
quality analysis, the quality can be measured just by counting the
places changes in the code occur in order to implement a functionality.
In an arbitrary MVC system when a \texttt{button} is dropped into the
VIEW part, the CONTROLLER needs some adjustment and maybe the MODEL,
too. Three places for adjustment is a reasonable easy task for the
brain. In the field of browser based development those principle is
often violated in the separated entities HTML, JS and CSS. A
\emph{browsernative microservice} aims for combining all of the
resources cohesively in a single place and file.

A \emph{component for the web}, or web component, is
\textbf{self-contained} which means it embodys all needed functionality
to get it's job done. Therefore it has a much better evolution mechanism
in the service contracts. Changing functionality won't break other
components. A component can progressively enhanced which guarantees
functionality throughout different versions.

Using libraries in web development is a common sense. But compared to
libraries, a component service offers multiple advantages. A library is
only loosely coupled to the implementation and therefore hard to track
in functionality. Changing a library may result in an unforeseen amount
of time fixing implementations. It is not unusual to see websites
embodying different versions of the same library (like with JQuery).
Another issue with libraries is dead code elimination which means the
process of removing code that is never going to be executed. New build
tools for the web, like Webpack 2 or Rollup offer this feature which
relies heavily on the static structure on ES6 modules.{[}6{]} Libraries
for the browsers are traditionally ``shipped'' as non static
\textbf{immediately-invoked functions}. Bootstrapping those is much
harder to archive as dead code may be part of the function itself. A
\emph{webnative microservices} should therefore embrace the high
cohesion principle and abandon most of the libraries in favor of static
ES6 module to lower the amount of dead code and to increase performance
especially at the first render.

Another advantage components have over libraries is the more explicit
interface.{[}4{]} While the functionality of a library needs
documentation, a component functionality is exposed via the components'
signature which comes in the fashion of an HTML element in
\emph{browsernative microservices}. For example, a button component can
consume attributes and becomes a primary buttons just by assignment

.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{<my-botton}\OtherTok{ isPrimary}\KeywordTok{>}\NormalTok{Primary}\KeywordTok{</my-botton>}
\end{Highlighting}
\end{Shaded}

@@

Emphasizing the human factor in microservices is a key feature.
Microservices are a product of real-world usage.{[}3, p. 1{]} Instead of
splitting team structures along the technology stack (UI Experts
-\textgreater{} Middleware -\textgreater{} Database) a microservice
approach model teams around \textbf{business capabilities}.{[}4{]}
Consequently every team is capable of planning, designing, implementing,
testing and maintaining their very own microservice. Along the
technology stack every member gains high competence about the
functionality of the service.

Real-world domains tend to be complex and multifaceted. To unfold their
complexity, domains can subdivided into \textbf{bounded contexts}.{[}3,
p. 31{]} For example, customer service is a business domain. Customers
again have different contexts depending on their demands. One context
can be sales, another context could be support. Every context makes
different assumptions about the underlying model. Each bounded context
draws an explicit interface where it decides what models to share with
other contexts.{[}3, p. 30{]} Each context can derived into multiple
microservices or, talking about \emph{Browsernative Microservices},
interfaces to the customers.

@@

For many companies working in the spheres of the internet the
client-side is highly important for their business. In fact, business
goals and capabilities can be derived from the front-end needs. The
state of the web is not only a story of numerous artifacts, it is also a
story of an highly fragmented market along devices, operating systems,
differing sizes and functionalities. Different devices again have
different assumptions about the technology stack. Splitting teams along
the stack results in an slow paced back and forth negotiation for every
change to be made. As browser technologies, design guidelines and
devices change in a fast paced manner it makes absolutely sense to shift
responsibility towards the teams altogether. A \emph{browsernative
microservice} embraces commitment over its whole lifecycle.

@@

Microservices for the browsers aim for side-effects like changing the UI
given to input parameters or emitting browser events. An isolated
microservice won't make too much sense after all. To ensure
communication between services, flexible yet powerful communication
channels must be established.

\emph{Smart endpoints and dumb pipes} is coined to the approach of
designing communication mostly decoupled and as cohesive as
possible.{[}4{]} In the analogy to the real world a message channel
should look like sending a letter: Two smart endpoints (sender and
receiver) and a mostly unified ``dumb'' letter format and channel. Each
endpoint owns it specific domain logic. In the browsernative context we
already emphasized the role of presentational components and containers.
The ``smartness'' of presentational components is altering the UI and
reacting on users input. The logic of a container follows the idea of an
filter in a Unix sense - receiving a request, applying logic as
appropriate and producing a response.{[}4{]}

Cited earlier in this paper microservices often rely on simple HTTP
request-response with resource API's and lightweight messaging as
communication protocols.{[}4{]} Those technologies highly accessible and
widely adopted and most probably usable for a browsernative service. In
the implementation part of this paper a reader will explore the
combination of RESTish protocols and a lightweight browser message bus
in action.

By focusing the service into clear defined business boundaries, it is
easier to define a smart API of the service which in our case will be a
set of attributes for the HTML element bound to the service.

@@

Microservices are separate entities and decentralization is therefore a
meta concept of microservices. Distributed service responsibility owned
by the team is one aspect. Gluing independent service bricks to a whole
distributed system talking over the network protocols another one.

This distributed nature allows teams to create their own technology
stack, tools and services designed in the spirit of language- and
platform independence and share their knowledge with other
parties.{[}4{]} In the recent years many big companies like Facebook,
Google, Netflix and others followed that spirit and published their
ideas and implementations open source. The previously mentioned ReactJS
for example is a brainchild of Facebook. In fact, many tools and
techniques are byproduct of vital interaction of concrete domain
problems and their implementations.

The spirit of freedom can't be applied universally to
\emph{Browsernative Microservices} as the browser and its underlying DOM
will be the limitation factor to a certain degree. Talking about the
browser, a reader might be tempted to narrowly thinking of the obvious
VIEW layer only - which is not true anymore. In the recent years the
browser engines grows to a kind of virtual machine: there are connectors
to build-in databases, multithreading support, an ever-growing JS
build-ins and even speech synthesis. So-called \emph{Progressive Web
Apps\footnote{\href{https://developers.google.com/web/progressive-web-apps}{Progressive
  Web Apps}}}, a bunch of criteria for building good browserapps, can
achieve a similar look and feel like native apps. Services like
NativeScript\footnote{\href{https://www.nativescript.org}{NativeScript}},
effectively compiling ``the web'' to native code, lower the boundary
between native and browser code even further.

@@

Another more real life decentralization aspect derives from the easiness
of deployment in a safe, sandboxed environment. Web components virtually
ship no overhead or require obscure build tools. This makes them ideal
candidates for sharing and open-sourcing. Webcomponents.org\footnote{\href{https://webcomponents.org}{Webcomponents.org}}
is a registry for ready-to-use components of every scale and purpose
where for even Google shares a lot of their material design elements.

@@

Data Management in Microservice follows the same modular philosophy like
the service implementation. As mentioned earlier different bounded
contexts make different assumptions of the underlying models. A
\emph{browsernative microservices} takes this idea even further and
expands it to the fragmented world of electronic devices. Decentralized
decisions about conceptual models demand for decentralized data storage
decisions.{[}4{]} Todays web architectures aims to leverage an
increasing amout of processing to the client to avoid time-consuming
roundtrips especially in mobiles networks.\footnote{Latency numbers:
  https://gist.github.com/jboner/2841832} Since network roundtrips are
costly it is a good advice to only query as much data as needed and
cache as much as possible.

@@

The simplified microservice example later in this paper assumes a
generic build-in API accompanied by build-in frontend components.
Instead of gluing frontend and backend together on runtime, the
microservice is designed holistically containing both front- and
backends. To reduce network calls especially for mobile devices it is a
good advice to cache data (f.e. in a global Service Worker). Revamping
offline capabilities even further data can be stored in a browser based
database like PouchDB. For the sake of simplicity data management won't
be explored into depth throughout this paper.

@@

Every check-in is handled as release candidate and can be independently
deployed by a fully automated deployment machinery.{[}8{]} Even though
this example is an opinionated perception it gives a sense of a mature
component build for the web. Due to an exhaustive amount of testing and
deployment tools for the browser an automated infrastructure shouldn't
be a problem.

@@

\emph{Browsernative Microservices} should be perceived as complementary
technology. Contrary to Angular, React and other frameworks they have a
strong interop with existing systems and can be used with them together.

@@

Depending on the context, some people argue for only two specs which
essentially make it possible to create a scoped component but not caring
too much on it's distribution{[}11{]}. Some people prefer the three
specs {[}12{]}, but the majority advocating the four specs variant,
which is listed on \sout{the official}
\href{http://webcomponents.org}{webcomponents.org} website. For the
purpose of this article, all four specs will be discussed briefly to
provide a rough understanding. It is not meant to cover all bits and
pieces.

@@

The current checkout can be found at
https://shop.polymer-project.org/checkout. Currently the checkout is a
single, 671 lines of code long Polymer component incorporation all
required fields for sign in, shipping, billing and summarizing the
order. In the spirit of microservices we will split up the component to
their smaller components. The shopping cart data is pulled out of a
local storage JSON entity set up previously by another custom-element.
Another in-memory opportunity is storing the data in the build-in
database IndexedDB.

@@

Yet already we see the simplicity around web components as they persue a
clear structure. Each of the child components should act independently
over other child nodes utilizing the loose coupling principle. Each
child ships all the HTML, CSS and JS code needed to fulfil its work
following the high cohesion principle. Each component may contain
different views to accommodate different bounded contexts resulting from
different devices. And last but not least, all of them communicate over
an unobstrusive message bus via the service root component
\texttt{shop-checkout}.

@@

Last but not least, there are \textbf{build-in components}, which is
every traditional HTMLElement. They are solely controlled and styled
from the outside and must therfore wrapped in presentationals and/or
containers.

Lets start the service description top-down with the \textbf{root
container}.

@@

Messages from child nodes will just forwarded to the worker and answers
passed back to the child nodes. Therefore, the basic microservice
communication looks like this

\begin{verbatim}
VIEW root           |               MODEL worker
                    |
Event  --------Event msg--------->    Msg
    |               |               Handler
    |               |                   |
Idle                |               Effekt
    |               |                   |
  Msg               |                   |
Handler <-----Action msg----------  Action                                      
                    |                   
                    |
\end{verbatim}

Effects are yielded by the asynchronous operation of messages and create
actions returned to sender. Effects can be created by external messages,
like subscription to an WebSocket, too. Effects may be created by
fetching additional resources from the server. The VIEW side of the root
element forwards all the messages and won't be bothered about the
content.

@@

\hypertarget{ref-Rauschmayer2015}{}
{[}6{]} D. A. Rauschmayer, ``Tree-shaking with webpack 2 and babel 6.''
2015 {[}Online{]}. Available:
\url{http://www.2ality.com/2015/12/webpack-tree-shaking.html}

>>>>

``A~\textbf{component}~is a unit of software that is independently
replaceable and upgradeable.'' {[}4{]} Components are the building
blocks of microservices. And microservices are the building blocks of
applications. Essentially the difference between microservices and
components is just the level of abstraction. Whether a concrete
microservice or a much more generic component, both share a similar set
of principles. Therefore this paper referring to both parts when talking
about \textbf{services}.

The first principle of services is the \textbf{loose coupling
principle}: changing and deploying one service shouldn't result in
changing other parts of the system.{[}3, p. 30{]}. Mutations or shadowed
variables, which is happening a lot in CSS, making it hard to keep
changes ought to only affect one place in the application. A
\emph{browsernative microservice} therefore pushing encapsulation and
avoiding variable mutations outside its scope as much as possible.
Practically, CSS will be scoped and JS fosters immutable JS entities and
avoids variables leaking into the global namespace.

The second principle of services is the \textbf{high cohesion
principle}: Whether designing a microservice or it's components we want
related behavior sit together, and unrelated behavior to sit
elsewhere.{[}3, p. 30{]} High cohesion can be expressed in a dynamic way
as the \emph{Single Responsibility Principle}: ``Gather together those
things that change for the same reason and separate those things that
change for different reasons.''{[}5{]} In a very quick and dirty code
quality analysis, the quality can be measured just by counting the
places changes in the code occur in order to implement a functionality.
An arbitrary threefold MVC system should require a maximum of three
changes to implement or change functionalities. The problem in browser
based development is not only the global paradigm which makes changes
deliberately unpredictable. The high cohesion principle is violated by
the traditional separation along the technological entities
\textbf{HTML, JS and CSS}. Understanding the relation of HTML markup and
another CSS file adds incidental complexity. Different approaches
emerged over the recent years to join the forces. Most notable the
dedicated frontend language \emph{elm}\footnote{\href{http://elm-lang.org/}{elm
  lang}} where model, view and controller sit in one place. A
browsernative service aims for a combinement of the web native trinity
HTML, JS and CSS.

Traditional web development relies on \textbf{libraries} to enhance the
service capabilities of the web platform. Compared to libraries, a
component service offers multiple advantages for building, deployment
and shipping. As expressed earlier a \emph{components for the web}, or
web component, is \textbf{self-contained} which means it embodies all
needed functionality to get it's job done. Therefore it has a much
better evolution mechanism in the service contracts. Changing
functionality won't break other services. A component can progressively
enhanced which guarantees functionality throughout different versions. A
library is only loosely coupled to the implementation and therefore hard
to track in functionality. Changing a library may result in an
unforeseen amount of time fixing implementations. It is not unusual to
see websites embodying different versions of the same library to
guarantee functionality but lowers page performance significantly.

Another issue where web components stand out is related to performance
and especially the critical first page rendering. Libraries for the
browsers are traditionally ``shipped'' as non static immediately-invoked
functions run immediately on page load. Following the \emph{Google RAIL}
model, a user-centric performance measurement, a page load ought be less
than 1 second to catch up user's attention.{[}6{]} There are many ways
for optimizing the critical first render but as a rule of thumb a
build-in web component might be always superior to libraries in terms of
first rendering. The fine-grained lifecycle methods which will be
described in the technical section of this paper give the developer far
reaching optimization opportunities.

The last argument in favor of components over libraries is the more
explicit interface.{[}4{]} While the functionality of a library needs
documentation to be accessible a component functionality is exposed via
the components' signature. The markup language HTML is by design
equipped with an expressive syntax and steering a web component using
attributes and values should be most straightforward even to
unexperienced web developers.

@@

Emphasizing the human factor in microservices is a key feature.
Microservices are a product of real-world usage.{[}3, p. 1{]} Instead of
splitting team structures along the technology stack (UI Experts
-\textgreater{} Middleware -\textgreater{} Database) a microservice
approach model teams around \textbf{business capabilities}.{[}4{]}
Consequently every team is capable of planning, designing, implementing,
testing and maintaining their very own microservice. Along the
technology stack every member gains high competence about the
functionality of the service architecture which can be critical for
evolving the service over its whole lifecycle.

Real-world domains tend to be complex and multifaceted. To unfold their
complexity, domains can subdivided into \textbf{bounded context}.{[}3,
p. 31{]} For example, customer service is a business domain but with
varying bounded context. One context can be sales, another context could
be support. Every context makes different assumptions about the
underlying model and draws an explicit interface where it decides what
to share with other contexts.{[}3, p. 30{]} Evaluating each bounded
context within each business domain will eventually shape the data
structures and services in question. This methodology can be iterated
over and over again. A sales service for example might be evaluated to
different sales context resulting in differing interfaces on devices
shaped by browsernative microservices.

@@

For many companies working in the spheres of the internet the client
side is highly important for their business. In fact, business goals and
capabilities can be derived from the front-end needs. The state of the
web is not only a story of numerous artifacts, it is also a story of an
highly fragmented market along devices, operating systems, differing
sizes and functionalities. Different devices again have different
assumptions about the technology stack. Splitting teams along the stack
results in an slow paced back and forth negotiation for every change to
be made. As browser technologies, design guidelines and devices change
frequently it makes absolutely sense to shift responsibility towards the
teams altogether. By focusing the service into clear defined business
boundaries, it is easier to define a smart API of the service.

@@

To ensure the microservice functionality among teams and different
services requires thoughtful decentralization. Emphasizing once more the
real-world capabilities of microservices a message channel architecture
can be derived from patterns known from traditional postal services. A
physical letter has only two smart endpoints entitled to read and
process the message while packaging likewise the connection itself is
maximized towards unification. \emph{Smart endpoints and dumb pipes} is
coined to the approach of designing communication mostly decoupled and
as cohesive as possible.{[}4{]} Applying this rules to the web platform
can result into building unified JSON message objects passed along
``dumb'' middleware components. A typical browser event comes close to
this definition and suits arguably well for in-memory communication
between web components and microservices.

Microservices heavily rely on simple HTTP request-response with resource
APIs and lightweight messaging.{[}4{]} REST APIs are technology agnostic
which aligns well with the idea of decentralization.{[}3, p. 247{]} The
advantage of this overall simple communication model is the suitability
for both frontend-backend likewise backend-backend communication. A
service therefore can evolve from an heavy backend with a lot of network
roundtrips to a leaner backend seamlessly. The browser build-in fetch
API which is essentially a HTTP request can be heavily incorporated into
a browsernative microservice to ensure communication between services.

@@

Microservices are separate entities and decentralization is important to
ensure autonomy. Distributed service responsibility owned by the team
promoting shared governance and technology agnostic communication
channels work towards a mostly decoupled service architecture.

This distributed nature allows teams to create their own technology
stack, tools and services designed in the spirit of language- and
platform independence and share their knowledge with other
parties.{[}4{]} In the recent years many big companies like Facebook,
Google, Netflix and others followed that spirit and published their
ideas and implementations open source. The previously mentioned ReactJS
for example is a product of Facebooks need to ensure a consistent
frontend experience. In fact, many tools and techniques are byproduct of
vital interaction of concrete domain problems and their implementations.

The spirit of freedom can't be applied universally to
\emph{browsernative microservices} as the browser and its underlying DOM
will be the limitation factor to a certain degree. Talking about the
browser, a reader might be tempted to narrowly thinking of the obvious
VIEW layer only - which is not true anymore. In the recent years the
browser engines grows to a kind of virtual machine: there are connectors
to build-in databases, multithreading support, an ever-growing JS
build-ins and even speech synthesis. So-called \emph{Progressive Web
Apps\footnote{\href{https://developers.google.com/web/progressive-web-apps}{Progressive
  Web Apps}}}, a bunch of criteria for building good browserapps, can
achieve a similar look and feel like native apps. Services like
NativeScript\footnote{\href{https://www.nativescript.org}{NativeScript}},
effectively compiling ``the web'' to native code, lower the boundary
between native and browser code even further.

@@

Another more real life decentralization aspect derives from the easiness
of deployment in a safe, sandboxed environment. Web components virtually
ship no overhead or require obscure build tools. This makes them ideal
candidates for sharing and open source publishing.
Webcomponents.org\footnote{\href{https://webcomponents.org}{Webcomponents.org}}
is a registry for ready-to-use components of every scale and purpose
where for even Google shares a lot of their material design elements.

@@

Data Management in a microservice follows the same modular philosophy
like the service implementation. As mentioned earlier different bounded
contexts make different assumptions of the underlying models. A
\emph{browsernative microservices} takes this idea even further and
expands it to the fragmented world of electronic devices. Decentralized
decisions about conceptual models demand for decentralized data storage
decisions.{[}4{]} Todays web architectures aims to leverage an
increasing amout of processing to the client to avoid time-consuming
roundtrips especially in mobiles networks.\footnote{Latency numbers:
  https://gist.github.com/jboner/2841832} Since network roundtrips are
costly it is a good advice to only query as much data as needed and
cache as much as possible. The build-in LocalStorage or its successor
IndexedDB are mature persistence technologies and libraries like
PouchDB\footnote{\href{https://pouchdb.com/}{PouchDB}} even offer
backend syncing out of the box.

@@

The simplified microservice example later in this paper assumes a
generic build-in API accompanied by build-in frontend components.
Instead of gluing frontend and backend together on runtime the
microservice is designed holistically containing both front- and
backends. For the sake of simplicity data management won't be explored
into depth throughout this paper.

@@

Every check-in is handled as release candidate and can be independently
deployed by a fully automated deployment machinery.{[}8{]} Even though
this example is an opinionated perception it gives a sense of a mature
component build for the web. Due to an exhaustive amount of testing and
deployment tools for JS an automated infrastructure shouldn't be a
problem.

@@

\emph{Browsernative microservices} should be perceived as complementary
technology. Contrary to Angular, React and other frameworks they have a
strong interop with existing systems and can be used with them together.

@@

Depending on the context, some people argue for only two specs which
essentially make it possible to create a scoped component but not caring
too much on it's distribution{[}11{]}. Some people prefer the three
specs {[}12{]}, but the majority advocating the four specs variant,
which is listed on the quasi-official
\href{http://webcomponents.org}{webcomponents.org} website. For the
purpose of this article, all four specs will be discussed briefly to
provide a rough understanding. It is not meant to cover all bits and
pieces.

@@

The current checkout can be found at
https://shop.polymer-project.org/checkout. Currently the checkout is a
single, 671 lines of code long Polymer component incorporation all
required fields for sign in, shipping, billing and summarizing the
order. In the spirit of microservices we will split up the microservice
into fine grained components. The shopping cart data is pulled out of a
local storage JSON entity set up previously by another custom-element.

@@

Yet already we see the simplicity arouse from web components as they
persue a clear structure. Each of the child components should act
independently over other child nodes utilizing the loose coupling
principle. Each child ships all the HTML, CSS and JS code needed to
fulfil its work following the high cohesion principle. Each component
may contain different views to accommodate different bounded contexts
resulting from different devices. And last but not least, all of them
communicate over an unobstrusive message bus via the service root
component \texttt{shop-checkout}.

@@

Last but not least, there are \textbf{build-in components}, which is
every build-in HTMLElement like the HTMLButtonElement. They are solely
controllable and styleable from the outside and must therefore wrapped
in presentational components and/or containers.

Lets start the service description top-down beginning with the
\textbf{root container} managing the overall service.

@@

The root container \texttt{shop-checkout} is basically just an
encapsulation layer in terms of service functionalities. Encapsulation
of CSS won't be necessary at this point as no styling is involved. On
initialization the \texttt{shop-checkout} might perform following tasks:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Importing child components with \textbf{HTML Imports}
\item
  Initialize the service in a logical top to bottom order like in the
  snippet above
\item
  Initializing a \textbf{model thread} with
  \texttt{new\ Worker(\textquotesingle{}checkout-model.js\textquotesingle{})};
\item
  Setting up a \texttt{postMessage} stub to dispatch messages to the
  model
\item
  Setting up an event listener for messages from the model
\item
  Setting up an event listener for \textbf{custom events} from the child
  nodes
\end{enumerate}

On runtime the root container acts mostly like a \textbf{message
dispatcher} implementing the microservice idea of smart endpoints and
dumb pipes.

Every custom child node is eligible aggregate subordinate events from
their children, buffer them and interact with the model trough via a
\textbf{dumb message system}. To talk to the model any child node can
dispatch a \textbf{unified event containing a unified message object}
like the following example illustrates:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{// Within every child node class}
\NormalTok{set }\AttributeTok{dispatch}\NormalTok{(msg) }\OperatorTok{\{}
  \KeywordTok{this}\NormalTok{.}\AttributeTok{dispatchEvent}\NormalTok{(}
    \KeywordTok{new} \AttributeTok{CustomEvent}\NormalTok{(}\StringTok{'checkout'}\OperatorTok{,} \OperatorTok{\{}
      \DataTypeTok{bubbles}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,}
      \DataTypeTok{detail}\OperatorTok{:} \NormalTok{msg}
    \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
  \NormalTok{)}\OperatorTok{;}
\OperatorTok{\}}
\end{Highlighting}
\end{Shaded}

Calling \texttt{this.dispatch(msg)} within a child node will trigger an
event bubbling upstream. The service root will implement a simple event
listener for \texttt{checkout} events. In a second step the service root
add information about the original target (like
\emph{tag-name\#id.className}) to the message so the message can be
processed in the model thread and eventually pushed back correctly to
the sender.

Summarizing the unidirectional model-view communication from the service
root perspective looks like following sketch:

\begin{verbatim}
VIEW thread         |           MODEL thread
                    |
Checkout ------Event Msg--------->    Msg
  Event             |                Handler
                    |                   |
                    |                Effect
                    |                   |
  Msg               |                Postmessage
Handler <-----Action Msg----------    Action
                    |                   
                    |
\end{verbatim}

Effects are yielded by the asynchronous operation of messages and create
actions returned to sender. Effects can be created by external messages,
like subscription to an WebSocket, too. Effects may be created with
additional resources from the server or syncing with local storage like
in the polymer-shop.

While the msg handlers are basic ``dumb'' switches the
\textbf{smartness} solely arouse from intelligent controllers processing
the message. Every component must implement a \textbf{unified socket} in
it's service contracts like a property on the custom element:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{// Within every child node class}
\NormalTok{set }\AttributeTok{receive}\NormalTok{(msg) }\OperatorTok{\{}
  \CommentTok{// do something smart}
\OperatorTok{\}}
\end{Highlighting}
\end{Shaded}

This communication flow is designed to \textbf{fail graceful} as simply
nothing happens if one or another switch is not defined. Furthermore it
incorporates lots of possibilities for \textbf{evolutionary design} as
child nodes can be loosely dropped in or teared apart. Middleware
container wrappers can be pulled in between child nodes and service root
to do some extra work and and even the message object can be enhanced
freely as long as it won't break with the message conventions. Changing
or enhancing any functionality requires only to touch the controllers in
the child nodes in question and the endpoint section at the model.
Furthermore, \textbf{infrastructure automation} might be achieved by
dynamically evaluating the mounted child nodes and registering the
``backend'' model switches on build time. Due to it's standardized
message system the components are testable within standardized tests.

To work in a larger scale, we might define a base component which has
the default message and socket properties in its guts and extend every
new element from this base class.

Another yet more powerful approach can be achieved by calculating the
new shape of the DOM in the worker directly and just patch the DOM in
the view layer.\footnote{A good example of this approach can be found at
  \href{http://www.pocketjavascript.com/blog/2015/11/23/introducing-pokedex-org}{Introducing
  Pokedex}} This approach is certainly more complicated but worth to
evaluate further more.

\subsubsection{Presentational
components}\label{presentational-components}

@@

\hypertarget{ref-Kearney2016}{}
{[}6{]} M. Kearney, ``Measure performance with the rail model.'' 2016
{[}Online{]}. Available:
\url{https://developers.google.com/web/fundamentals/performance/rail}
